---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fable.tbats

<!-- badges: start -->
[![R-CMD-check](https://github.com/JSzitas/fable.tbats/workflows/R-CMD-check/badge.svg)](https://github.com/JSzitas/fable.tbats/actions)
[![Codecov test coverage](https://codecov.io/gh/JSzitas/autodep/branch/main/graph/badge.svg)](https://codecov.io/gh/JSzitas/autodep?branch=main)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-green.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![CRAN status](https://www.r-pkg.org/badges/version/fable.tbats)](https://CRAN.R-project.org/package=fable.tbats)
<!-- badges: end -->

fable.tbats is a wrapper around the implementation of **tbats** from the [forecast](https://github.com/robjhyndman/forecast) package. 
This integrates seamlessly, and can be used just like any model in **fable**: 

```{r, message=FALSE, warning=FALSE}
library(tsibbledata)
library(fable)
library(fable.tbats)
library(dplyr)

# fit models to the pelt dataset until 1930:
train <- pelt %>% 
  filter(Year < 1930)
test <- pelt %>% 
  filter(Year >= 1930)

models <- train %>% 
  model( ets = ETS(Lynx),
         bats = BATS(Lynx),
         tbats = TBATS(Lynx)
         ) 
# generate forecasts on the test set
forecasts <- forecast(models, test)
# visualize
autoplot(forecasts, pelt)
```

Similarly, accuracy calculation works: 

```{r, }
train_accuracies <- accuracies <- accuracy(models)
knitr::kable(train_accuracies)
```

```{r}
test_accuracies <- accuracy(forecasts, test)
knitr::kable(test_accuracies)
```

As does refitting:

```{r}
models <- refit( models, pelt )
```

Of the functionality available in the **forecast** package, only **forecast::tbats.components()** 
is missing. 

## Performance note 

Fitting bats/tbats to a few long series can (potentially) be slower using this wrapper 
than using the forecast package. This is due to the fact that the internal tbats/bats algorithm 
always executes sequentially (i.e. with **use.parallel = FALSE** ) to prevent issues with nested parallelism 
(as the **fabletools::model** function is taken to be responsible for handling parallelization). 

This should never be a problem on many time series, but does lead to a significant slow-down if you are only 
modelling a single/few time series. Nonetheless, in those cases the [fasster](https://github.com/tidyverts/fasster) package might be much better suited 
for your use case anyways. 

## Installation

Currently only: 

``` r
pak::pkg_install("JSzitas/fable.tbats")
```


