---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# fable.tbats

<!-- badges: start -->
[![R-CMD-check](https://github.com/JSzitas/fable.tbats/workflows/R-CMD-check/badge.svg)](https://github.com/JSzitas/fable.tbats/actions)
[![Codecov test coverage](https://codecov.io/gh/JSzitas/fable.tbats/branch/main/graph/badge.svg)](https://codecov.io/gh/JSzitas/fable.tbats?branch=main)
[![Lifecycle: stable](https://img.shields.io/badge/lifecycle-stable-green.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
[![CRAN status](https://www.r-pkg.org/badges/version/fable.tbats)](https://CRAN.R-project.org/package=fable.tbats)
<!-- badges: end -->

**fable.tbats** provides the BATS and TBATS models of De Livera, Hyndman and Snyder (2011)
for **fable**. The fitting is implemented in C++ inside the package (a single, dependency-free
header, `src/tbats/tbats.h`, designed in `src/tbats/DESIGN.md`); it began as a wrapper around the
[forecast](https://github.com/robjhyndman/forecast) package's implementation, against which it
is verified. 

## Installation

``` r
pak::pkg_install("JSzitas/fable.tbats")
```

## Usage 

Used just like any model in **fable**: 

```{r, message=FALSE, warning=FALSE}
library(tsibbledata)
library(fable)
library(fable.tbats)
library(dplyr)

# fit models to the pelt dataset until 1930:
train <- pelt %>% 
  filter(Year < 1930)
test <- pelt %>% 
  filter(Year >= 1930)

models <- train %>% 
  model( ets = ETS(Lynx),
         bats = BATS(Lynx),
         tbats = TBATS(Lynx)
         ) 
# generate forecasts on the test set
forecasts <- forecast(models, test)
# visualize
autoplot(forecasts, pelt)
```

Similarly, accuracy calculation works: 

```{r, }
train_accuracies <- accuracy(models)
knitr::kable(train_accuracies)
```

```{r}
test_accuracies <- accuracy(forecasts, test)
knitr::kable(test_accuracies)
```

## Simulation

Simulated future sample paths are available through **generate()**. Innovations are
drawn from a normal distribution with the fitted variance by default, or resampled from
the innovation residuals with **bootstrap = TRUE**:

```{r}
paths <- models %>%
  select(bats, tbats) %>%
  generate(h = "10 years", times = 20, seed = 1)
paths
```

```{r, message=FALSE}
library(ggplot2)
paths %>%
  ggplot(aes(x = Year, y = .sim, group = .rep)) +
  geom_line(alpha = 0.4) +
  geom_line(aes(y = Lynx, group = NULL), data = pelt, colour = "steelblue") +
  facet_wrap(~ .model, ncol = 1) +
  labs(y = "Lynx")
```

```{r}
bootstrapped <- models %>%
  select(tbats) %>%
  generate(h = "10 years", times = 20, bootstrap = TRUE, seed = 1)
```

Note that **residuals()** returns the innovation residuals of the state space model by
default, which are on the Box-Cox transformed scale when a Box-Cox transformation is
used. Residuals on the scale of the data are available with **type = "response"**.

## Refitting

Refitting to new data works as well:

```{r}
models <- refit( models, pelt )
```

### A note on refitting

`refit()` deliberately departs from the forecast package. With the default
`reestimate = FALSE` the estimated parameters are kept and only the seed states
(the state vector before the first observation) are re-estimated for the new
series, because the stored seed states describe the origin of the training
series and nothing else. The forecast package reuses them unchanged, which is
only right when the new series starts at the same origin. With
`reestimate = TRUE` the full model search is repeated on the new data.

## Components

The level, the slope and one seasonal effect per period are available as a dable,
on the Box-Cox transformed scale when the model uses a transformation:

```{r}
models %>%
  select(tbats) %>%
  components()
```


## Performance note 

The model search runs sequentially; **fabletools::model** is responsible for parallelising
across series. Like for like (sequential on both sides), the C++ implementation is 1.5 to 3
times faster than the forecast package's on series from 72 to 5000 observations, with an
equal or lower AIC in every case; `experiments/benchmark_port.R` reproduces the comparison
against the implementation this package carried up to version 0.5.0. 
